Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-07 doc/report/msc-report.tex
doc/report/msc-report.tex: * changed tense to passive in Implementation sections. * minor cosmetic changes.

HBase has a Thrift server which serves as the interface to languages other
than Java--for instance Python.

@@

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map (in the MapReduce sense) and letting Zohmg run that map over the data
set. The user's map is called \textit{usermap} from here on. In order to let Zohmg
know what output to expect from the mapper, the user also defines the
dimensions and units of his data. After all, the user most likely knows his
own data better than anyone else.

A standard MapReduce map emits key-value pairs. Not so in the case of the
Zohmg mapper: the usermap emits triples (that is, a tuple of three values)
consisting of a timestamp, a list of dimensions and their respective values,
and a list of measurements. In effect, the dimensions and the timestamp
specify a point in n-space and the measurements represent the observations
made at that point.

Going back to the running example of apache access logs, a usermap might look
something like this:

@@

The actual parsing of the log is abstracted; it is the emitting of
dimensionality data I want to showcase in this example. We assume that
\texttt{parse\_apache\_log()} accepts a single line from the log file and
returns an object with the appropriate fields, such as useragent ant
statuscode.

For every line of the log file, we emit a point in n-space (the dimensions and
the timestamp) together with a count of pageviews and bytes; one pageview per
request and the number of bytes served in the request. It is likely that you
will want to add business logic here, perhaps only counting requests from
useragents that aren not spiders and that resulted in a status code of 200.

The last line \textit{yields} the tuple. This is a python construct called a
generator that lets the map emit any number of tuples (or none!) for a single
invocation.

@@

\begin{verbatim}
  zohmg import mappers/apache.py /data/weblogs/2009/05/10
\end{verbatim}

This will instruct zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the \texttt{/data/weblogs/2009/05/10} directory
on HDFS. The emitted output is duly collected and projections are carefully
stored in the data store.

@@

When reasoning about dimensions, the issue of cardinality will come up sooner
or later.

The cardinality of a dimension is the number of distinct values of the
dimension. For example, the dimension of \textit{logged in} has a cardinality
of two since \textit{logged in} is a boolean variable; the user is either
logged in or not. The \textit{country} dimension has a much higher
cardinality: there are some 300 countries in the world. Certainly, there could
be dimension whose cardinality is infinite: the \textit{path} dimension in the
web log example, for example, could in theory have a limitless number of
values.

The number of possible combinations of dimensions of your data---the number
of points in n-space---is directly related to the cardinality of each
dimension. It is likely that your data is of such a nature that it is
physically impossible to store information about all dimension combinations.
This is why we ask you, the user, to define what projections you are
interested in: storing aggregates for the slices of the cube that interest you
is both a time and space save, and in many cases the only feasible solution.

@@

Since we wish to store projections rather than all combinations of all
dimensions, we need to mangle the data slightly before sending it down the
pipe to the underlying data store.

Dumbo expects a key-value pair from the mapper. Hadoop will collect all keys
and give the reducer a key and an iterator of values. If we are to store
projections, we need to encode both the desired dimensions and the unit in
question in the key and emit the value of the unit in the \textit{value} field.
This way we can use a simple SumReducer to sum the values for each unit at any
point in n-space and be done.

@@

The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a json-serialized map of the form
{"column-family:qualifier" : value}.

The OutputReader interprets the json and persists the data. Et voila!

>>>>

HBase has a Thrift server which serves as the interface to languages other
than Java---for instance Python.

@@

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map function (in the MapReduce sense) and letting Zohmg run that map function
over the data set. The user's map function is called \textit{usermap function}
from here on. In order to let Zohmg know what output to expect from the mapper,
the user also defines the dimensions and units of his data. After all, the user
most likely knows his own data better than anyone else.

A standard MapReduce map function emits key-value pairs. Not so in the case of
the Zohmg mapper: the usermap function emits triples (that is, a tuple of three
values) consisting of a timestamp, a list of dimensions and their respective
values, and a list of measurements. In effect, the dimensions and the timestamp
specify a point in n-space and the measurements represent the observations made
at that point.

Going back to the running example of apache access logs, a usermap function
might look something like this:

@@

The actual parsing of the log is abstracted; in the above example the
emitting of dimensionality is showcased. It is assumed that
\texttt{parse\_apache\_log()} accepts a single line from the log file and
returns an object with the appropriate fields, such as useragent and
statuscode.

For every line of the log file, a point in n-space (the dimensions and the
timestamp) is emitted together with a count of pageviews and bytes; one
pageview per request and the number of bytes served in the request. It is
possible to add business logic here, perhaps only counting requests from
useragents that are not web spiders and that resulted in a HTTP 200 status
code.

The last line \textit{yields} the tuple. This is a Python construct called a
generator that lets the map function emit any number of tuples (or none!)
for a single invocation.

@@

\begin{verbatim}
   zohmg import mappers/apache.py \
   /data/weblogs/2009/05/10
\end{verbatim}

This will instruct zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the directory
\texttt{/data/weblogs/2009/05/10} on HDFS. The emitted output is duly
collected and projections are carefully stored in the data store.

@@

TODO?: When reasoning about dimensions, the issue of cardinality will come
up sooner or later.

The cardinality of a dimension is the number of distinct values of the
dimension, see it as number of members in a set where the dimension is
viewed as a set. For example, the dimension of \textit{logged in} has a
cardinality of two since \textit{logged in} is a boolean variable; the user
is either logged in or not. The \textit{country} dimension has a much
higher cardinality: there are some 300 countries in the world. Certainly,
there could be dimensions whose cardinality is infinite: the \textit{path}
dimension in the web log example, for example, could in theory have a
limitless number of values.

The number of possible combinations of dimensions of the data---the number
of points in n-space---is directly related to the cardinality of each
dimension. It is likely that the user's data is of such a nature that it is
physically impossible to store information about all dimension
combinations.  This is why the user is asked to define what projections are
interesting: storing aggregates for the slices of the cube that is
interesting is both a time and space save, and in many cases the only
feasible solution.

@@

Since the wish is to store projections rather than all combinations of all
dimensions, the data needs to be mangled slightly before sending it down
the pipe to the underlying data store.

Dumbo expects a key-value pair from the mapper. Hadoop will collect all keys
and give the reducer a key and an iterator of values.

If projections are to be stored, both the desired dimensions and the unit
in question needs to be encoded in the key, and the value of the unit
emitted in the \textit{value} field.

By doing this it is possible to use a simple SumReducer to sum the values
for each unit at any point in n-space.

@@

The reducer needs to communicate to the OutputReader under what rowkey,
column-family and qualifier to store the computed sum. It does this by
emitting the row key as the key and a JSON-serialized map of the form
\texttt{\{"column-family:qualifier" : value\}}.

The OutputReader interprets the JSON and persists the data. Et voila!
